Distributional Trees and Forests

### 2017-06-22 pkg/disttree/inst/useR2017/slides.Rnw
changed loaded simlist and deleted other rda-files
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@1477 255f4249-2263-4417-b1ea-8498779e27a7

  if(!(any(dt.par, df.par, g.par, all.par))){
    par(mar=c(5.1,4.1,4.1,3.1))
    plot(y = sp$y, x = sp$x, type = "p", col="grey", main = "True parameters", xaxt='n', yaxt='n', xlab = "", ylab = "")
    lines(x = sp$x, y = sp$true.mu, type = "l", col = 'black')
    lines(x = sp$x, y = sp$true.mu + sp$true.sigma, type = "l", lty = 2, col = 'black')
    lines(x = sp$x, y = sp$true.mu - sp$true.sigma, type = "l", lty = 2, col = 'black')
    legend('topleft', c(TeX('$\\mu$'), TeX('$\\mu \\pm \\sigma$')), 
           col = c('black','grey'), lty = 1, bty = 'n')
  }

@@

  # gamlss
  if(g.par){
    par(mar=c(5.1,4.1,4.1,3.1))
    plot(y = sp$y, x = sp$x, type = "p", col="grey",  main = "gamlss", xaxt='n', yaxt='n', xlab = "", ylab = "")
    lines(x = sp$x, y = sp$true.mu, type = "l", col = 'black')
    lines(x = sp$x, y = sp$true.mu + sp$true.sigma, type = "l", col = 'grey')
    lines(x = sp$x, y = sp$true.mu - sp$true.sigma, type = "l", col = 'grey')
    lines(x = sp$x, y = sp$fitted.mu.g, type = "l", col = pal["gamlss"], lwd = 1.5)
    lines(x = sp$x, y = sp$fitted.mu.g + sp$fitted.sigma.g, type = "l", col = 'lightblue')
    lines(x = sp$x, y = sp$fitted.mu.g - sp$fitted.sigma.g, type = "l", col = 'lightblue')
    legend('topleft', c(TeX('$\\mu$'), TeX('$\\mu \\pm \\sigma$'), 
                        TeX('$\\hat{\\mu}$'), TeX('$\\hat{\\mu} \\pm \\hat{\\sigma}$')),  
           col = c('black','grey',pal["gamlss"], 'lightblue'), lty = 1, bty = 'n')
  }

@@

<<preliminaries3, echo=FALSE, results=hide>>=
load("simlist20170621.rda")

@@

  if(type == "exp"){
    #dev.off() 
    rmse <- cbind(simlist$av.rmse.exp.true.dt, simlist$av.rmse.exp.true.df, simlist$av.rmse.exp.true.g,
                  simlist$av.rmse.exp.obs.dt, simlist$av.rmse.exp.obs.df, simlist$av.rmse.exp.obs.g)
    colnames(rmse) <- c("dt.true", "df.true", "g.true", "dt.obs", "df.obs", "g.obs")
    ylim <- c(min(rmse), max(rmse))
    par(mar=c(5.1,4.1,4.1,3.1))
    plot(x = simlist$x.axis, y = rmse[,"dt.true"], type = "l", col = pal["tree"], ylim = ylim,
         xlab = "", ylab = "", main = "disttree vs distforest vs gamlss", xaxt='n', yaxt='n', lwd = 2)
    lines(x = simlist$x.axis, y = rmse[,"dt.obs"], type = "l", lty = 2, col = pal["tree"], lwd = 2)
    lines(x = simlist$x.axis, y = rmse[,"df.true"], type = "l", col = pal["forest"], lwd = 2)
    lines(x = simlist$x.axis, y = rmse[,"df.obs"], type = "l", lty = 2, col = pal["forest"], lwd = 2)
    lines(x = simlist$x.axis, y = rmse[,"g.true"], type = "l", col = pal["gamlss"], lwd = 2)
    lines(x = simlist$x.axis, y = rmse[,"g.obs"], type = "l", lty = 2, col = pal["gamlss"], lwd = 2)
    #legend('left', c("disttree", "distforest", "gamlss"), col = c(pal["tree"], pal["forest"], pal["gamlss"]), lty = 1)
    mtext(text= "smooth", side = 1, line = 1, adj = 0)
    mtext(text= "steep", side = 1, line = 1, adj = 1)
    mtext(text= TeX('$\\kappa$'), side = 1, line = 1)
    mtext(text= "RMSE", side = 2, line = 1)
    mtext(text= "disttree", side = 2, col = pal["tree"], las = 1, line = 0.2, padj = 6.5)
    mtext(text= "distforest", side = 2, col = pal["forest"], las = 1, line = 0.2, padj = 13)
    mtext(text= "gamlss", side = 2, col = pal["gamlss"], las = 1, line = 0.3, padj = 16)

@@

  if(type == "par"){
    rmse <- cbind(simlist$av.rmse.mu.dt, simlist$av.rmse.mu.df, simlist$av.rmse.mu.g, 
                  simlist$av.rmse.sigma.dt, simlist$av.rmse.sigma.df, simlist$av.rmse.sigma.g)
    colnames(rmse) <- c("mu.dt", "mu.df", "mu.g", "sigma.dt", "sigma.df", "sigma.g")
    ylim <- c(min(rmse), max(rmse))
    par(mar=c(5.1,4.1,4.1,3.1))
    plot(x = simlist$x.axis, y = rmse[,"mu.dt"], type = "l", col = pal["tree"], ylim = ylim,
         xlab = "", ylab = "", main = "disttree vs distforest vs gamlss", xaxt='n', yaxt='n', lwd = 2)
    lines(x = simlist$x.axis, y = rmse[,"sigma.dt"], type = "l", lty = 2, col = pal["tree"], lwd = 2)
    lines(x = simlist$x.axis, y = rmse[,"mu.df"], type = "l", col = pal["forest"], lwd = 2)
    lines(x = simlist$x.axis, y = rmse[,"sigma.df"], type = "l", lty = 2, col = pal["forest"], lwd = 2)
    lines(x = simlist$x.axis, y = rmse[,"mu.g"], type = "l", col = pal["gamlss"], lwd = 2)
    lines(x = simlist$x.axis, y = rmse[,"sigma.g"], type = "l", lty = 2, col = pal["gamlss"], lwd = 2)
    #legend('topleft', c("disttree", "distforest", "gamlss"), col = c(pal["tree"], pal["forest"],pal["gamlss"]), lty = 1)
    mtext(text= "smooth", side = 1, line = 1, adj = 0)
    mtext(text= "steep", side = 1, line = 1, adj = 1)
    mtext(text= TeX('$\\kappa$'), side = 1, line = 1)
    mtext(text= "RMSE", side = 2, line = 1)
    mtext(text= "disttree", side = 2, col = pal["tree"], las = 1, line = 0.2, padj = -4)
    mtext(text= "distforest", side = 2, col = pal["forest"], las = 1, line = 0.2, padj = 8)
    mtext(text= "gamlss", side = 2, col = pal["gamlss"], las = 1, line = 0.3, padj = 14)

@@

# plot loglikelihood
plot_ll <- function(simlist){
  ll <- cbind(simlist$av.loglik.dt, simlist$av.loglik.df, simlist$av.loglik.g)
  colnames(ll) <- c("dt", "df","g")
  ylim <- c(min(ll), max(ll))
  par(mar=c(5.1,4.1,4.1,3.1))
  plot(x = simlist$x.axis, y = ll[,"dt"], type = "l", col = pal["tree"], ylim = ylim,
       xlab = "", ylab = "", main = "disttree vs distforest vs gamlss", xaxt='n', yaxt='n', lwd = 2)
  lines(x = simlist$x.axis, y = ll[,"df"], type = "l", col = pal["forest"], lwd = 2)
  lines(x = simlist$x.axis, y = ll[,"g"], type = "l", col = pal["gamlss"], lwd = 2)
  #legend('topright', c("disttree", "distforest", "gamlss"), col = c(pal["tree"], pal["forest"],pal["gamlss"]), lty = 1)
  mtext(text= "smooth", side = 1, line = 1, adj = 0)
  mtext(text= "steep", side = 1, line = 1, adj = 1)
  mtext(text= TeX('$\\kappa$'), side = 1, line = 1)
  mtext(text= "Log-Likelihood", side = 2, line = 1)
  mtext(text= "disttree", side = 2, col = pal["tree"], las = 1, line = 0.2, padj = 12)
  mtext(text= "distforest", side = 2, col = pal["forest"], las = 1, line = 0.2, padj = -8.5)
  mtext(text= "gamlss", side = 2, col = pal["gamlss"], las = 1, line = 0.3, padj = -16)

>>>>

  if(!(any(dt.par, df.par, g.par, all.par))){
    par(mar=c(5.1,4.1,4.1,3.1))
    plot(y = sp$y, x = sp$x, type = "p", col="grey", main = "True parameters", xaxt='n', yaxt='n', xlab = "", ylab = "")
    lines(x = sp$x, y = sp$true.mu, type = "l", col = 'black')
    lines(x = sp$x, y = sp$true.mu + sp$true.sigma, type = "l", lty = 2, col = 'black')
    lines(x = sp$x, y = sp$true.mu - sp$true.sigma, type = "l", lty = 2, col = 'black')
    legend('topleft', c(TeX('$\\mu$'), TeX('$\\mu \\pm \\sigma$')), 
           col = c('black','black'), lty = c(1,2,2), bty = 'n')
  }

@@

  # gamlss
  if(g.par){
    par(mar=c(5.1,4.1,4.1,3.1))
    plot(y = sp$y, x = sp$x, type = "p", col="grey",  main = "gamlss", xaxt='n', yaxt='n', xlab = "", ylab = "")
    lines(x = sp$x, y = sp$true.mu, type = "l", col = 'black')
    lines(x = sp$x, y = sp$true.mu + sp$true.sigma, type = "l", col = 'grey')
    lines(x = sp$x, y = sp$true.mu - sp$true.sigma, type = "l", col = 'grey')
    lines(x = sp$x, y = sp$fitted.mu.g, type = "l", col = pal["gamlss"], lwd = 2)
    lines(x = sp$x, y = sp$fitted.mu.g + sp$fitted.sigma.g, type = "l", col = 'lightblue')
    lines(x = sp$x, y = sp$fitted.mu.g - sp$fitted.sigma.g, type = "l", col = 'lightblue')
    legend('topleft', c(TeX('$\\mu$'), TeX('$\\mu \\pm \\sigma$'), 
                        TeX('$\\hat{\\mu}$'), TeX('$\\hat{\\mu} \\pm \\hat{\\sigma}$')),  
           col = c('black','grey',pal["gamlss"], 'lightblue'), lty = 1, bty = 'n')
  }

@@

<<preliminaries3, echo=FALSE, results=hide>>=
load("simlist20170621b.rda")

@@

  if(type == "exp"){
    #dev.off() 
    rmse <- cbind(simlist$av.rmse.exp.true.dt, simlist$av.rmse.exp.true.df, simlist$av.rmse.exp.true.g,
                  simlist$av.rmse.exp.obs.dt, simlist$av.rmse.exp.obs.df, simlist$av.rmse.exp.obs.g)
    colnames(rmse) <- c("dt.true", "df.true", "g.true", "dt.obs", "df.obs", "g.obs")
    ylim <- c(min(rmse), max(rmse))
    par(mar=c(5.1,4.1,4.1,3.1))
    plot(x = simlist$x.axis, y = rmse[,"dt.true"], type = "l", col = pal["tree"], ylim = ylim,
         xlab = "", ylab = "", main = "disttree vs distforest vs gamlss", xaxt='n', yaxt='n', lwd = 2)
    lines(x = simlist$x.axis, y = rmse[,"dt.obs"], type = "l", lty = 2, col = pal["tree"], lwd = 2)
    lines(x = simlist$x.axis, y = rmse[,"df.true"], type = "l", col = pal["forest"], lwd = 2)
    lines(x = simlist$x.axis, y = rmse[,"df.obs"], type = "l", lty = 2, col = pal["forest"], lwd = 2)
    lines(x = simlist$x.axis, y = rmse[,"g.true"], type = "l", col = pal["gamlss"], lwd = 2)
    lines(x = simlist$x.axis, y = rmse[,"g.obs"], type = "l", lty = 2, col = pal["gamlss"], lwd = 2)
    #legend('left', c("disttree", "distforest", "gamlss"), col = c(pal["tree"], pal["forest"], pal["gamlss"]), lty = 1)
    mtext(text= "smooth", side = 1, line = 1, adj = 0)
    mtext(text= "steep", side = 1, line = 1, adj = 1)
    mtext(text= TeX('$\\kappa$'), side = 1, line = 1)
    mtext(text= "RMSE", side = 2, line = 1)
    mtext(text= "disttree", side = 2, col = pal["tree"], las = 1, line = 0.2, padj = 5)
    mtext(text= "distforest", side = 2, col = pal["forest"], las = 1, line = 0.2, padj = 13)
    mtext(text= "gamlss", side = 2, col = pal["gamlss"], las = 1, line = 0.3, padj = 16.5)

@@

  if(type == "par"){
    rmse <- cbind(simlist$av.rmse.mu.dt, simlist$av.rmse.mu.df, simlist$av.rmse.mu.g, 
                  simlist$av.rmse.sigma.dt, simlist$av.rmse.sigma.df, simlist$av.rmse.sigma.g)
    colnames(rmse) <- c("mu.dt", "mu.df", "mu.g", "sigma.dt", "sigma.df", "sigma.g")
    ylim <- c(min(rmse), max(rmse))
    par(mar=c(5.1,4.1,4.1,3.1))
    plot(x = simlist$x.axis, y = rmse[,"mu.dt"], type = "l", col = pal["tree"], ylim = ylim,
         xlab = "", ylab = "", main = "disttree vs distforest vs gamlss", xaxt='n', yaxt='n', lwd = 2)
    lines(x = simlist$x.axis, y = rmse[,"sigma.dt"], type = "l", lty = 2, col = pal["tree"], lwd = 2)
    lines(x = simlist$x.axis, y = rmse[,"mu.df"], type = "l", col = pal["forest"], lwd = 2)
    lines(x = simlist$x.axis, y = rmse[,"sigma.df"], type = "l", lty = 2, col = pal["forest"], lwd = 2)
    lines(x = simlist$x.axis, y = rmse[,"mu.g"], type = "l", col = pal["gamlss"], lwd = 2)
    lines(x = simlist$x.axis, y = rmse[,"sigma.g"], type = "l", lty = 2, col = pal["gamlss"], lwd = 2)
    #legend('topleft', c("disttree", "distforest", "gamlss"), col = c(pal["tree"], pal["forest"],pal["gamlss"]), lty = 1)
    mtext(text= "smooth", side = 1, line = 1, adj = 0)
    mtext(text= "steep", side = 1, line = 1, adj = 1)
    mtext(text= TeX('$\\kappa$'), side = 1, line = 1)
    mtext(text= "RMSE", side = 2, line = 1)
    mtext(text= "disttree", side = 2, col = pal["tree"], las = 1, line = 0.2, padj = -7)
    mtext(text= "distforest", side = 2, col = pal["forest"], las = 1, line = 0.2, padj = 8)
    mtext(text= "gamlss", side = 2, col = pal["gamlss"], las = 1, line = 0.3, padj = 14)

@@

# plot loglikelihood
plot_ll <- function(simlist){
  ll <- cbind(simlist$av.loglik.dt, simlist$av.loglik.df, simlist$av.loglik.g)
  colnames(ll) <- c("dt", "df","g")
  ylim <- c(min(ll), max(ll))
  par(mar=c(5.1,4.1,4.1,3.1))
  plot(x = simlist$x.axis, y = ll[,"dt"], type = "l", col = pal["tree"], ylim = ylim,
       xlab = "", ylab = "", main = "disttree vs distforest vs gamlss", xaxt='n', yaxt='n', lwd = 2)
  lines(x = simlist$x.axis, y = ll[,"df"], type = "l", col = pal["forest"], lwd = 2)
  lines(x = simlist$x.axis, y = ll[,"g"], type = "l", col = pal["gamlss"], lwd = 2)
  #legend('topright', c("disttree", "distforest", "gamlss"), col = c(pal["tree"], pal["forest"],pal["gamlss"]), lty = 1)
  mtext(text= "smooth", side = 1, line = 1, adj = 0)
  mtext(text= "steep", side = 1, line = 1, adj = 1)
  mtext(text= TeX('$\\kappa$'), side = 1, line = 1)
  mtext(text= "Log-Likelihood", side = 2, line = 1)
  mtext(text= "disttree", side = 2, col = pal["tree"], las = 1, line = 0.2, padj = 16)
  mtext(text= "distforest", side = 2, col = pal["forest"], las = 1, line = 0.2, padj = -8)
  mtext(text= "gamlss", side = 2, col = pal["gamlss"], las = 1, line = 0.3, padj = -16)
